### 2015-04-25 paper/paper.tex
Add paper outline

\documentclass{article} % For LaTeX2e
\usepackage{nips15submit_e,times}
\usepackage{hyperref}
\usepackage{url}
%\documentstyle[nips14submit_09,times,art10]{article} % For LaTeX 2.09

\title{
%TODO: we did a thing
Deep learning on psychological arousal classification}

\author{
Yueran Yuan, Jeffrey Zhang, David Rayson, Kai-min Chang\\
Language Technologies Institute\\
Carnegie Mellon University\\
Pittsburgh, PA 15213\\
\texttt{\{yueranyuan, jzhang94, davidr92, kaimin.chang\}@gmail.com}
}

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}

\nipsfinalcopy % Uncomment for camera-ready version, comment for submission version!

\begin{document}

\maketitle

\begin{abstract}
%TODO: hi so like there's this problem, we tried to do this. it kinda works not really
\end{abstract}

\section*{Introduction}
%TODO: so like there's this problem right? we wanted to solve it and we think our approach would be a good idea. in section 1 we'll discuss the things people have tried before. in section 2 we'll discuss the data we had. in section 3 we'll discuss our approach. in section 4 we'll discuss why we're legit scientists. in section 5 we'll discuss what results we got. in the conclusion we'll wax philosophical.
%-the problem in general
%-what we did in general and why
%-what the sections are

\section{Prior work}
%TODO: so like our thing has never actually been done before. actually, this guy did this one thing but it really sucks. so we were like obviously we could do better with technology x.
%-EEG
% -what is EEG
% -EEG patterns (psych stuff/neuro stuff)
% -Dataset (brief)
%-Deep learning
% -what is deep learning
% -deep learning + EEG

\section{Data}
%TODO: we have big data, oooh. okay maybe not that big but it still takes forever to load 
%-Dataset (details)
% -processing
%-Features
% -raw
% -dtft
%  -windows
% -wavelets
% -other

\section{Techniques}
%TODO: we thought technology x would be a great idea because it has these advantages over the sucky way that one guy was doing it. so we wrote some code and did it. we wrote the code. :o
%-Deep learning
% -deep nets
% -dropout
% -autoencoder
% -denoting autoencoder
% -convolutional neural nets
% -parameterized rectifiers
% -batchnorm (plus all that other stuff)
%-Ensemble
% -random forest
% -classifier fusion

\section{Experiment}
%TODO: we're not making up numbers here yo.
%-Cross-validation
%-Parameter search
%-Evaluation
% -baseline classifiers
% -auc, acc, best-cutoff
%-Alternative evaluations
%-Ranking channels for importance

\section{Results}
%TODO: yeah so um our thing sucks like the other guy's stuff. but it sucks less. so yaya
%-the results

\section*{Conclusion}
%TODO: look at what we did. we think we can do better next time. but yeah we'll need money for ramen so gib pls
%-refer back to the prev work
%-future work

\subsubsection*{Acknowledgments}
%TODO: thanks mom

\subsubsection*{References}
%TODO: [1] awesome paper by awesome guy

\end{document}